---
title: "Comparisons"
---

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(chorddiag)
library(htmlwidgets)
library(igraph)
library(readr)
library(tidygraph)
library(tidyverse)

```

## Comparing Past vs. Present Conditions

Using LANDFIRE’s BpS products, we explore two different ways to visualize past vs. current vegetation patterns.

-   First we present **changes in broad ecosystem types** using an interactive comparison diagram. To present these broad ecosystem trends more effectively, we combined classifications of Existing Vegetation Types (EVT) into broader categories.Developed EVTs of high, medium, and low intensity, and Quarries-Strip Mines-Gravel Pits EVTs were combined into one "Developed" EVT. Exotic Herbaceous and Exotic Tree-shrub EVTs were combined into one "Exotics" EVT.

-   Second we compare **amounts of succession classes** (past and present) for the most prevalent ecosystems.

## Comparing Broad Vegetation Trends

### Summary

NEED TO FILL THIS IN

*Note: number presented when hovering equals acres.*

```{r chord, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# read in data
chord_df <- read_csv("data/bps2evt_chord.csv")
#view(histFireGVchord)

#convert to matrix
matrix_df <- as.matrix(as_adjacency_matrix(as_tbl_graph(chord_df),attr = "ACRES"))

#clean up matrix (could be cleaner!)
matrix_df = subset(matrix_df, select = -c(1:3))

matrix_df <- matrix_df[-c(4:10),]

#make a custom color pallet #eb4034 (redish) #b0af9e(grey)

# ORIGINAL
groupColors <- c( "#1d4220", # conifer 
                 "#fc9d03", # grassland
                 "#56bf5f", # hardwood
                 "#397d3f", # hardwood-conifer 
                 "#7db7c7", # riparian 
                 "#f5e942", # cur ag
                 "#1d4220", # cur conifer
                 "#397d3f", # cur hdw-con
                 "#b0af9e", # developed
                 "#eb4034", # exotics
                 "#fc9d03", # grassland
                 "#56bf5f", # hardwood
                 "#7db7c7",
                 "#6e4f1e"# shrubland
                 

)



#make chord diagram
chord<-chorddiag(data = matrix_df,
                 type = "bipartite",
                 groupColors = groupColors,
                 groupnamePadding = 10,
                 groupPadding = 3,
                 groupnameFontsize = 12 ,
                 showTicks = FALSE,
                 margin=150,
                 tooltipGroupConnector = "    &#x25B6;    ",
                 chordedgeColor = "#363533"
)
chord 

#save then print to have white background
htmlwidgets::saveWidget(chord,
                        "chord.html",
                        background = "white",
                        selfcontained = TRUE
                        )
```

<iframe src="chord.html" height="720" width="720" style="border: 1px solid #464646;" allowfullscreen allow="autoplay" data-external=".5">

</iframe>

<br>

## Succession classes for three most dominant BpSs

NEED TO FILL THIS IN




```{r scls chart, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

aoi_ref_cur <- read.csv("data/bpsScls2.csv")

## wrangle data, get top groups (BpSs)

bps_scls_top <- aoi_ref_cur %>%
  group_by(model_code) %>%
  mutate(total_count = ifelse(total_count == 0, max(total_count), total_count)) %>%
  arrange(desc(total_count))  %>%
  ungroup() %>%
  dplyr::filter(dense_rank(desc(total_count)) < 4) %>%
  dplyr::select(c("bps_name", "ref_label",  "current_percent", "ref_percent")) %>%
  pivot_longer(
    cols = c(`ref_percent`, `current_percent`), 
    names_to = "ref_cur", 
    values_to = "percent"
    )


# order classes
bps_scls_top$ref_label <- factor(bps_scls_top$ref_label, levels = c(
  "Developed",
  "Agriculture",
  "UE",
  "UN",
  "E",
  "D",
  "C",
  "B",
  "A"))

sclasplot <-
  ggplot(bps_scls_top, aes(fill = factor(ref_cur), y = percent, x = ref_label)) + 
  geom_col(width = 0.8, position = position_dodge()) +
  coord_flip() +
  facet_grid(. ~BpS) +
  scale_x_discrete(limits = (levels(bps_scls_top$ref_label))) +
  labs(
    title = "Succession Classes past and present",
    subtitle = "Top BpSs selected for illustration. Not all succession classes present in all BpSs",
    caption = "Data from landfire.gov.",
    x = "",
    y = "Percent") +
  theme_minimal(base_size = 12) +
  theme(plot.caption = element_text(hjust = 0, face = "italic"), #Default is hjust=1
        plot.title.position = "plot", #NEW parameter. Apply for subtitle too.
        plot.caption.position =  "plot") +
  scale_fill_manual(values = c("#3d4740", "#32a852" ), # present (grey), historical (green)
                    name = " ", 
                    labels = c("Present",
                               "Past")) +
  facet_wrap(~bps_name, nrow(3),labeller = labeller(bps_name = label_wrap_gen())) +
    theme(panel.spacing = unit(.05, "lines"),
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        strip.background = element_rect(color = "black", size = 1))

sclasplot

```
